Design and Implementation of Probabilistic Programming Language Anglican

### 2016-08-05 paper/paper.tex
minor

Operational semantics of Anglican queries is different from that
of Clojure code, therefore queries must be called through
inference algorithms, rather than `directly'.  The
\texttt{anglican.inference} namespace declares the polymorphic
function \texttt{infer} using Clojure's multimethod mechanism.
This function accepts an Anglican query and
returns a lazy sequence of weighted samples from the
distribution defined by the query.  When inference
is performed on an Anglican query, the query is run by a
particular inference algorithm. Inference algorithms must
provide an implementation for \texttt{infer}, as well as
override the polymorphic function \texttt{checkpoint} (defined
as a multimethod) so as to handle \texttt{sample} and
\texttt{observe} in an algorithm-specific manner, as well
as to construct an appropriate result on the 
termination of a probabilistic program.

@@

The Anglican language is a subset of Clojure\footnote{It would be
possible to support almost full Clojure by expanding all macros
in the Anglican source code, however in Clojure, unlike in
Scheme~\cite{SDF+10}, or Common Lisp~\cite{PC94}, the result
of macro-expansion of derived special forms is not well
specified and implementation specific.}.  Anglican queries
are defined within \texttt{defquery} as shown in the previous section,
using \texttt{if}, \texttt{when}, \texttt{cond}, \texttt{case}, 
\texttt{let}, \texttt{and}, \texttt{or}, \texttt{fn} forms;
other forms of Clojure may be in the future but are not now. 
In \texttt{let} bindings and \texttt{fn} argument lists,
Clojure's pattern matching mechanism for vector data type
(called vector destructuring) is supported. Also, compound 
literals for vectors, hash maps, and
sets are supported just like in Clojure.

\subsection{Tail recursion}

Clojure provides special constructs \texttt{loop} and \texttt{recur}
for writing tail-recursive programs. Such constructs are not necessary
in Anglican, because Anglican programs are CPS-converted and do not
use stacks; no recursive call in Anglican can lead to stack overflow.
In Anglican, recursive calls to functions should be used instead of
\texttt{recur}. However, \texttt{loop}/\texttt{recur} is provided in Anglican
for convenience as a way to express loops. \texttt{recur} outside of \texttt{loop} 
will lead to unpredictable behaviour and hard-to-catch errors.

\subsection{Core library}

@@

>>>>

Operational semantics of Anglican queries is different from that
of Clojure code, therefore queries must be called through
inference algorithms, rather than `directly'.  The
\texttt{anglican.inference} namespace declares the polymorphic
function \texttt{infer} using Clojure's multimethod mechanism.
This function accepts an Anglican query and
returns a lazy sequence of weighted samples from the
distribution defined by the query.  When inference
is performed on an Anglican query, the query is run by a
particular inference algorithm. Inference algorithms must
provide an implementation for \texttt{infer}, as well as
override the polymorphic function \texttt{checkpoint} (defined
as a multimethod) so as to handle \texttt{sample} and
\texttt{observe} in an algorithm-specific manner and
to construct an appropriate result on the 
termination of a probabilistic program.

@@

The Anglican language is a subset of Clojure\footnote{It would be
possible to support almost full Clojure by expanding all macros
in the Anglican source code, however in Clojure, unlike in
Scheme~\cite{SDF+10}, or Common Lisp~\cite{PC94}, the result
of macro-expansion of derived special forms is not well
specified and implementation specific.}.  Anglican queries
are defined within \texttt{defquery} as shown in the previous section,
using \texttt{if}, \texttt{when}, \texttt{cond}, \texttt{case}, 
\texttt{let}, \texttt{and}, \texttt{or}, \texttt{fn} forms;
other forms of Clojure may be supported in the future but are not now. 
In \texttt{let} bindings and \texttt{fn} argument lists,
Clojure's pattern matching mechanism for vector data type
(called vector destructuring) is supported. Also, compound 
literals for vectors, hash maps, and
sets are supported just like in Clojure.

\subsection{Core Library}

@@

\subsection{Tail Call}

Clojure provides special forms \texttt{loop} and \texttt{recur}
for writing tail-recursive programs. Such forms are not necessary
in Anglican, because Anglican programs are CPS-converted and do not
use stacks. For instance, no recursive call in Anglican can lead to stack overflow.
In fact, in Anglican, it is recommended to use recursive calls to functions instead of
\texttt{recur}. However, \texttt{loop}/\texttt{recur} is provided in Anglican
for convenience as a way to express loops. \texttt{recur} outside of \texttt{loop} 
will lead to unpredictable behaviour and hard-to-catch errors.
